---
title: "Introduction"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
knitr::opts_chunk$set(echo = FALSE)
```


## Basic Operations

### Simple Mathematics

*R can be used to perform calculations.*


```{r simple_math, exercise=TRUE, exercise.lines = 7}
2 + 3  

4 * 5

1 / 3

log(10)
```


### Exercises

Write the R code to add 3 plus 4:

```{r simple_math_test, exercise=TRUE, exercise.lines = 3}

```

```{r simple_math_test-solution}
3 + 4
```


## Objects


### Single Values

*You can assign values to a name, this is known as creating an object. The `<-` operator assigns the value to the object name.*  

*Calling the object name will return the value.*

```{r simple_object, exercise=TRUE, exercise.lines = 4}
my_object <- 4

my_object
```


### Vector of values

*You can also assign a set (or vector) of values to a name.*  

*A vector of values is written using `c(var1, var2, var3)`.*

```{r simple_vector, exercise=TRUE, exercise.lines = 4}
my_vector <- c(1, 4, 6)

my_vector
```


### Exercises

*Vectors can contain other types of values such as character strings.*

Write the R code to create a vector containing the two words Hello and World:

```{r vector_test, exercise=TRUE, exercise.lines = 3}

```

```{r vector_test-solution}
character_vector <- c("Hello", "World")
character_vector
```


## Functions


*R has lots of built in functions to perform various operations*  
*There are also many packages available which have other useful functions*  

```{r prepare-functions}
my_vector <- c(1, 4, 6)
```

```{r sample_functions, exercise=TRUE, exercise.setup = "prepare-functions"}
mean(my_vector)
max(my_vector)
seq(1, 10)

one_to_ten <- seq(1, 10)
length(one_to_ten)
```

### Helpfiles

*There are help files for all functions. The help files will explain what the function should do and what arguments should/can be given. Help files can be accesed using ?*

```{r helpfile, exercise=TRUE}
?mean
```

*The helpfile explains the mean function and the available arguments. There is an optional argument `na.rm`, this removes NA values. A NA value is a blank entry*

What happens when we try to calculate the mean of a set of variables which contain an NA entry?

An additional arguments can fix this, use the helpfile to fix the code below?

```{r function_arguments, exercise=TRUE, exercise.lines = 5}
my_vector2 <- c(3, 4, 5, NA)
my_vector2

mean(my_vector2)

```

```{r function_arguments-solution}
mean(my_vector2, na.rm = TRUE)
```

